---
title: "A Comprehensive Guide to Fractional Regression Models with fracreg"
author: "Sulman Olieko Owili"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Comprehensive Guide to Fractional Regression Models with fracreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The **`fracreg`** package is the most comprehensive tool available in R for estimating, diagnosing, and interpreting fractional regression models. It supports:
- Univariate fractional models (1P)
- Hurdle two-part fractional models (2P)
- Double-inflated three-part fractional models (3P)
- Fractional models with panel data (`fracregpd`)
- Fractional models with endogenous covariates or heteroskedasticity (`fracreghet`)

This vignette will walk you through simulated examples to demonstrate how to use each of these powerful estimators.

```{r setup}
library(fracreg)
```

## 1. Cross-Sectional Fractional Models (`fracreg`)

The core `fracreg()` function is designed to handle univariate models where the dependent variable is bounded between 0 and 1 inclusive ($0 \le y \le 1$). We will simulate some data to demonstrate its usage.

### Simulating the Data
```{r data-sim}
set.seed(123)
N <- 1000
x1 <- rnorm(N)
x2 <- runif(N)

# Generating a fractional dependent variable with inflation at 0 and 1
XB <- -0.5 + 0.8 * x1 + 1.2 * x2 + rnorm(N)
y_latent <- exp(XB) / (1 + exp(XB))

y <- y_latent
# Inflate at boundaries
y[y_latent < 0.2] <- 0
y[y_latent > 0.8] <- 1

data <- data.frame(y = y, x1 = x1, x2 = x2)
head(data)
```

### The Standard 1-Part Model (1P)
If you assume a single process governs the entire distribution of `y`, you can use the standard one-part model.

```{r one-part}
model_1p <- fracreg(
  y = data$y, 
  x = cbind(x1 = data$x1, x2 = data$x2), 
  type = "1P", 
  linkfrac = "logit"
)
summary(model_1p)
```

### Partial Effects (1P)
To interpret the coefficients in terms of their marginal impact on $y$, you can compute the partial effects using the `fracreg.pe()` command:
```{r pe-1p}
pe_1p <- fracreg.pe(model_1p)
summary(pe_1p)
```

### The Two-Part Hurdle Model (2P)
Often, fractional data (like participation rates or test scores) have a large cluster of zeros. The two-part model estimates the binary "hurdle" ($y>0$) and the continuous fractional part ($0 < y \le 1$) sequentially, then combines them.

```{r two-part}
# We use inflation=0 to indicate we are modeling the mass at zero
model_2p <- fracreg(
  y = data$y, 
  x = cbind(x1 = data$x1, x2 = data$x2), 
  type = "2P", 
  inflation = 0,
  linkbin = "logit", 
  linkfrac = "logit"
)
summary(model_2p)

# Calculate partial effects for the combined two-part model
pe_2p <- fracreg.pe(model_2p)
summary(pe_2p)
```

### The Three-Part Double-Inflated Model (3P)
When your data has substantial clustering at **both** boundaries (0 and 1) simultaneously, the 3-part model decomposes the process into three stages:
1. The probability of $y > 0$
2. The probability of $y = 1$ given $y > 0$
3. The fractional outcome $0 < y < 1$

```{r three-part}
# Linkbin takes a vector of two links for the two binary hurdles
model_3p <- fracreg(
  y = data$y, 
  x = cbind(x1 = data$x1, x2 = data$x2), 
  type = "3P", 
  linkbin = c("logit", "logit"), 
  linkfrac = "logit"
)
summary(model_3p)

# The analytical delta method is natively supported for the 3P model!
pe_3p <- fracreg.pe(model_3p)
summary(pe_3p)
```

---

## 2. Hypothesis Testing and Specification Diagnostics

`fracreg` includes state-of-the-art specification tests to check whether your functional form and link functions are adequate.

### Generalized Goodness-Of-Functional-Form (GGOFF)
```{r ggoff}
# Perform the GGOFF test on the 1P model
# Tests if the logit link is appropriate
fracreg.ggoff(model_1p, version = "LM")
```

### RESET Test
```{r reset}
# Standard RESET test for functional form misspecification
fracreg.reset(model_1p)
```

### P-Test for Non-Nested Models
You can test non-nested models (e.g., comparing a `cloglog` link against a `logit` link) using `fracreg.ptest()`.

```{r ptest}
model_1p_clog <- fracreg(y = data$y, x = cbind(x1 = data$x1, x2 = data$x2), type = "1P", linkfrac = "cloglog")
# fracreg.ptest(model_1p, model_1p_clog) # Uncomment to run non-nested test
```

---

## 3. Endogeneity & Heteroskedasticity (`fracreghet`)

When you suspect that one of your covariates is endogenous, or that the variance is heteroskedastic, you can use `fracreghet`. It natively supports instrumental variables (IV) through a Control Function approach or GMM estimation.

### Control Function Approach
```{r endogeneity}
# Simulating an endogenous variable (var.endog) and an instrument (z)
z <- rnorm(N)
u <- 0.5 * z + rnorm(N)
var.endog <- 0.8 * z + u
y_endog <- exp(0.5 * x1 + 1.2 * var.endog + u) / (1 + exp(0.5 * x1 + 1.2 * var.endog + u))

# Estimate QML with control function for cross-sectional data
model_het <- fracreghet(
  y = y_endog, 
  x = cbind(x1, var.endog), 
  z = cbind(x1, z), 
  var.endog = var.endog,
  type = "QMLxv", 
  link = "logit"
)
```

---

## 4. Panel Data Fractional Models (`fracregpd`)

For longitudinal or panel data where unobserved heterogeneity is a concern, `fracregpd()` provides fixed-T panel estimators, including Correlated Random Effects (CRE).

```{r panel-data}
# Simulating Panel Data
id <- rep(1:200, each = 5)
time <- rep(1:5, times = 200)
x_panel <- rnorm(1000)
# Unobserved individual effect (CRE)
c_i <- rep(rnorm(200), each = 5) 
y_panel <- exp(x_panel + c_i) / (1 + exp(x_panel + c_i))

# Estimate a Correlated Random Effects (CRE) Model
model_pd <- fracregpd(
  id = id, 
  time = time, 
  y = y_panel, 
  x = cbind(x_panel = x_panel), 
  type = "QMLcre", 
  link = "probit"
)
```

## Conclusion

With `fracreg`, `fracreghet`, and `fracregpd`, you have a complete, mathematically robust toolkit for modeling fractional responses bounded between [0,1], regardless of inflation, endogeneity, or unobserved panel effects.
