---
title: "A Comprehensive Guide to Fractional Response Regressions with fracreg"
author: "Sulman Olieko Owili"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Comprehensive Guide to Fractional Response Regressions with fracreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = TRUE
)

local({
  hook_output <- knitr::knit_hooks$get("output")
  knitr::knit_hooks$set(output = function(x, options) {
    paste0("\n<details><summary>Toggle to see the output</summary>\n\n",
           hook_output(x, options),
           "\n</details>\n")
  })
})

library(fracreg)
```

## Introduction

The **`fracreg`** package is the most comprehensive tool available in R for estimating, diagnosing, and interpreting fractional response models. Originally focused on standard cross-sectional models, it has been substantially expanded to provide a complete suite of fractional response methodologies. The package supports:

- **Cross-Sectional Models (`fracreg`)**: Includes one-part models (1P) for standard $[0, 1]$ fractions, two-part hurdle models (2P) for fractions with boundaries at 0 or 1, and three-part double-inflated models (3P) for inflation at both boundaries.
- **Panel Data Models (`fracregpd`)**: Utilises Correlated Random Effects (CRE), Quasi-Maximum Likelihood (QML), and Generalised Method of Moments (GMM) approaches for longitudinal data.
- **Endogeneity Correction (`fracreghet`)**: Corrects for endogenous covariates using Control Function and IV-GMM approaches.
- **Fractional Multinomial Logit (`fracregmlogit`)**: Models multivariate fractional outcomes where multiple fractions sum to 1.
- **Fractional Ridge Regression (`fracregridge`)**: Incorporates L2 regularization dynamically controlled by the fraction of the unregularized coefficient vector length.

This vignette walks you through comprehensive empirical examples (using the built-in 401(k) and spending datasets) and simulated examples for each estimator. We provide detailed explanations of why each model is used and how to interpret the associated outputs.

```{r load-package}
library(fracreg)
```

## 1. Data Description and Preparation

We use the built-in `fracreg_k401k` dataset to demonstrate univariate fractional models. This is the canonical firm-level 401(k) plan participation data used in **Papke and Wooldridge (1996)** (*"Econometric methods for fractional response variables with an application to 401(k) plan participation rates"*, Journal of Applied Econometrics). The dataset contains 1,534 observations of 401(k) plans.

The primary variables we use are:
- `prate`: The plan participation rate (the fraction of eligible employees who are active participants). It strictly falls in the $[0, 1]$ interval and serves as our dependent variable ($y$). 
- `mrate`: The firm's matching rate (the firm's contribution per $1 of employee contribution). 
- `age`: The age of the 401(k) plan. 
- `totemp`: Total number of employees at the firm. 
- `sole`: A binary indicator equal to 1 if the 401(k) plan is the sole retirement plan offered by the firm.

---

## 2. Cross-Sectional Fractional Models (`fracreg`)

The core `fracreg()` function is designed for univariate models where the dependent variable is bounded between 0 and 1 inclusive ($0 \le y \le 1$). 

### 2.1 Empirical 401(k) Examples

We begin by modeling the `prate` using the available firm characteristics.

#### 1P Model (One-Part)
The simplest model is the 1P model, which assumes a single quasi-likelihood process governs the entire distribution of the fractional response. Here, we estimate a 1P model using the standard `logit` link function. We also demonstrate how to obtain Odds Ratios (exponentiated coefficients) with 99% confidence intervals, which can sometimes provide more intuitive multiplicative interpretations of the covariates' impacts.

```{r fracreg-empirical-1p}
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
# 1P Model using logit link
mod_1p <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(mod_1p)

# 1P Model reporting odds ratios and 99% confidence intervals
mod_1p_or <- fracreg(y, X, type="1P", linkfrac="logit", or=TRUE, level=0.99)
summary(mod_1p_or)
```

#### 2P and 3P Models
Often, fractional data contains a large cluster (inflation) of observations exactly at 0 or 1. A 2P (two-part) model separates the decision to participate at the boundary from the magnitude of participation. Below, we estimate a 2P model that specifically models inflation at 1 (using `inflation=1`). We then simulate some 0s to demonstrate a 3P (three-part) model, which models inflations at both 0 and 1 using two binary link functions and one fractional link function.

```{r fracreg-empirical-multip}
# 2P Model (modelling mass at 1) 
mod_2p <- fracreg(y, X, type="2P", inflation=1, linkbin="logit", linkfrac="logit") 
summary(mod_2p)
 
# 3P Model (injecting artificial 0s for demonstration) 
y_3p <- y; y_3p[1:50] <- 0 
mod_3p <- fracreg(y_3p, X, type="3P", linkbin=c("logit","logit"), linkfrac="logit") 
summary(mod_3p)
```

### 2.2 Simulated Cross-Sectional Examples

To illustrate boundary effects more clearly, we simulate a fractional variable that is artificially bounded and inflated at 0 and 1. We then estimate the fractional response using different model types.

```{r fracreg-simulated}
set.seed(123) 
N <- 1000 
x1 <- rnorm(N) 
x2 <- runif(N) 
 
# Generate a fractional dependent variable with inflation at 0 and 1 
XB <- -0.5 + 0.8 * x1 + 1.2 * x2 + rnorm(N) 
y_latent <- exp(XB) / (1 + exp(XB)) 
 
y_sim <- y_latent 
# Inflate at boundaries 
y_sim[y_latent < 0.2] <- 0 
y_sim[y_latent > 0.8] <- 1 
X_sim <- cbind(x1 = x1, x2 = x2) 
 
# Standard 1P logit model
mod_sim_1p <- fracreg(y_sim, X_sim, type="1P", linkfrac="logit") 
summary(mod_sim_1p)
```

If we are only interested in estimating one part of a two-part model, we can isolate it. Here, we estimate only the binary hurdle component (`type="2Pbin"`) and then only the fractional continuous component (`type="2Pfrac"`).

```{r fracreg-simulated-2p-parts}
# 2P Model: Binary component only (is y > 0?)
mod_sim_2pbin <- fracreg(y_sim, X_sim, type="2Pbin", inflation=0, linkbin="logit") 
summary(mod_sim_2pbin)
 
# 2P Model: Fractional component only (given y > 0) using probit link
mod_sim_2pfrac <- fracreg(y_sim, X_sim, type="2Pfrac", inflation=0, linkfrac="probit") 
summary(mod_sim_2pfrac)
```

Finally, we estimate both components simultaneously (a full 2P model) and a 3P double-inflated model. Note that different link functions (e.g., `cloglog`, `probit`, `logit`) can be mixed within the same multi-part model.

```{r fracreg-simulated-3p}
# Full Two-Part Model (cloglog binary link + logit fractional link)
mod_sim_2p_full <- fracreg(y_sim, X_sim, type="2P", inflation=0, linkbin="cloglog", linkfrac="logit") 
summary(mod_sim_2p_full)
 
# Three-part double-inflated model (y has both 0s and 1s) 
mod_sim_3p <- fracreg(y_sim, X_sim, type="3P", linkbin=c("logit","probit"), linkfrac="logit") 
summary(mod_sim_3p)
```

### 2.3 Partial Effects (`fracreg.pe`)

Because fractional models are inherently non-linear, raw coefficients cannot be interpreted as constant marginal effects. The `fracreg.pe()` function computes the actual marginal effects. You can request Average Partial Effects (APE) across all observations or Conditional Partial Effects (CPE) at specific covariate values (e.g., at the median).

```{r fracreg-pe}
# Computing APE for the empirical 401(k) 1P logit model
m <- fracreg(y, X, type="1P", linkfrac="logit", table=FALSE) 
pe_res <- fracreg.pe(m) 
summary(pe_res)
 
# Computing APE for a simulated 2P Model
mod_2p_sim <- fracreg(y_sim, X_sim, linkbin="logit", linkfrac="probit", type="2P", inf=1, table=FALSE) 
pe_2p <- fracreg.pe(mod_2p_sim) 
summary(pe_2p)
 
# Computing CPE (Conditional Partial Effect) for X2, evaluating all covariates at their median values 
mod_2pfrac <- fracreg(y_sim, X_sim, linkfrac="logit", type="2Pfrac", inf=1, table=FALSE) 
pe_cpe <- fracreg.pe(mod_2pfrac, APE=FALSE, CPE=TRUE, at="median", which.x="x2") 
summary(pe_cpe)
```

---

## 3. Hypothesis Testing and Specification Diagnostics

`fracreg` includes state-of-the-art specification tests to validate your model's functional form and link function assumptions, helping you ensure the statistical validity of your estimates.

### 3.1 Generalised Goodness-Of-Functional-Form (`fracreg.ggoff`)

The GGOFF test checks if the chosen link function (e.g., `logit`, `probit`) is appropriate for the data. A significant test statistic suggests that the link function may be misspecified and an alternative should be explored.

```{r fracreg-ggoff}
# Testing the logit specification of the empirical 401(k) model
m <- fracreg(y, X, type="1P", linkfrac="logit", table=FALSE) 
ggoff_res <- fracreg.ggoff(m) 
summary(ggoff_res)
 
# Testing a binary probit specification for a simulated 2P hurdle using the Likelihood Ratio (LR) test
mod_bin <- fracreg(y_sim, X_sim, linkbin="probit", type="2Pbin", inf=1, table=FALSE) 
ggoff_bin <- fracreg.ggoff(mod_bin, "LR") 
summary(ggoff_bin)
```

### 3.2 RESET Test (`fracreg.reset`)

The RESET test detects general functional form misspecification (such as omitted variables or incorrect non-linearities) by testing whether fitted powers of the linear predictor have significant explanatory power when added back into the model.

```{r fracreg-reset}
# RESET test on the empirical 401(k) 1P model
reset_res <- fracreg.reset(m) 
summary(reset_res)
 
# Using LM and Wald versions of RESET with 2nd and 3rd order fitted powers on simulated data
mod_sim <- fracreg(y_sim, X_sim, linkfrac="logit", table=FALSE) 
reset_sim <- fracreg.reset(mod_sim, 2:3, c("Wald","LM")) 
summary(reset_sim)
```

### 3.3 P-Test for Non-Nested Models (`fracreg.ptest`)

When you need to choose between two non-nested model specifications (for example, a `logit` link vs. a `probit` or `cloglog` link), the P-test compares them directly to see if one statistically dominates the other.

```{r fracreg-ptest}
# Comparing logit vs. probit 1P models empirically
m1 <- fracreg(y, X, type="1P", linkfrac="logit", table=FALSE) 
m2 <- fracreg(y, X, type="1P", linkfrac="probit", table=FALSE) 
ptest_res <- fracreg.ptest(m1, m2) 
summary(ptest_res)
 
# Comparing a 1P logit model against a 2P model (logit binary + probit fractional) using a Wald test
res1 <- fracreg(y_sim, X_sim, linkfrac="logit", table=FALSE) 
res2 <- fracreg(y_sim, X_sim, linkbin="logit", linkfrac="probit", type="2P", inf=1, table=FALSE) 
ptest_mix <- fracreg.ptest(res1, res2, "Wald") 
summary(ptest_mix)
```

---

## 4. Endogeneity & Heteroscedasticity (`fracreghet`)

When a covariate is endogenous (correlated with the error term) or heteroscedasticity is present, standard maximum likelihood estimation becomes inconsistent. `fracreghet()` provides Instrumental Variable (IV) corrections using Control Function (CF) and Generalised Method of Moments (GMM) approaches.

### 4.1 Estimating Endogenous Models
In the empirical example below, we suspect `mrate` (matching rate) might be endogenous. We use `age` and `ltotemp` (log of total employees) as instruments. We estimate the model using a Quasi-Maximum Likelihood control function approach (`QMLxv`).

```{r fracreghet}
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X_het <- cbind(mrate = fracreg_k401k$mrate, ltotemp = fracreg_k401k$ltotemp)
 
# fracreghet estimators do not allow exact 1s or 0s. 
# In practice, boundary values are strictly shifted inward by a minute amount.
y_adj <- y
y_adj[y_adj == 1] <- 0.999

# Instrument mrate using age and ltotemp
Z_emp <- cbind(age = fracreg_k401k$age, ltotemp = fracreg_k401k$ltotemp) 

# Estimate using QML Control Function approach
mod_het <- fracreghet(y_adj, X_het, Z_emp, var.endog = X_het[, "mrate"], type="QMLxv", link="logit") 
summary(mod_het)
```

We can also simulate a scenario where `var.endog` is explicitly endogenous due to a shared unobserved term `u`. We then correct it using various GMM approaches.

```{r fracreghet-simulated}
set.seed(123) 
N <- 1000 
x1 <- rnorm(N) 
z1 <- rnorm(N) 
u <- 0.5 * z1 + rnorm(N) 

# Create endogenous variable correlated with u
var.endog <- 0.8 * z1 + u 
y_endog <- exp(0.5 * x1 + 1.2 * var.endog + u) / (1 + exp(0.5 * x1 + 1.2 * var.endog + u)) 
 
y_endog[y_endog <= 0] <- 0.01 
y_endog[y_endog >= 1] <- 0.99 
 
X_sim <- cbind(x1 = x1, var.endog = var.endog) 
Z_sim <- cbind(x1 = x1, z1 = z1) 
 
# GMMz estimator: Uses instruments directly without requiring a linear reduced form for the endogenous variable
mod_gmmz <- fracreghet(y = y_endog, x = X_sim, z = Z_sim, type = "GMMz", link = "logit") 
summary(mod_gmmz)
 
# QMLxv estimator: Control function approach assuming a linear reduced form for the endogenous variable
mod_qmlxv <- fracreghet(y = y_endog, x = X_sim, z = Z_sim, var.endog = var.endog, type = "QMLxv", link = "logit") 
summary(mod_qmlxv)
```

### 4.2 Partial Effects (`fracreghet.pe`)

Partial effects can also be calculated for endogenous models. When using a control function approach (`QMLxv`), `fracreghet.pe` accurately integrates over the estimated first-stage residuals.

```{r fracreghet-pe}
# Computing partial effects for the empirical endogenous model
res_emp <- fracreghet(y_adj, X_het, Z_emp, var.endog = X_het[, "mrate"], type="QMLxv", link="logit", table=FALSE) 
pe_emp <- fracreghet.pe(res_emp, which.x="mrate")
summary(pe_emp)
```

---

## 5. Panel Data Fractional Models (`fracregpd`)

When you observe the same entities over multiple time periods, unobserved individual heterogeneity must be accounted for. `fracregpd()` supports fixed-T panel estimators, including Correlated Random Effects (CRE) and panel GMM.

Below, we simulate a panel dataset where individual specific effects (`c_i`) influence the outcome.

```{r fracregpd}
set.seed(123) 
N <- 100 
T_periods <- 5 
id <- rep(1:N, each = T_periods) 
time <- rep(1:T_periods, times = N) 
x_panel <- rnorm(N * T_periods) 
 
# Unobserved individual effect (CRE) 
c_i <- rep(rnorm(N), each = T_periods)  
y_panel <- exp(x_panel + c_i) / (1 + exp(x_panel + c_i)) 
X_panel <- cbind(x_panel = x_panel) 
 
# Estimate a Correlated Random Effects (CRE) Model 
# CRE explicitly models the individual effect by including time-averages of covariates
mod_cre <- fracregpd(id=id, time=time, y=y_panel, x=X_panel, type="QMLcre", link="probit") 
summary(mod_cre)
 
# GMMww estimator: Panel GMM allowing for lagged covariates and robust standard errors
mod_gmmww <- fracregpd(id=id, time=time, y=y_panel, x=X_panel, lags=TRUE, type="GMMww", var.type="robust") 
summary(mod_gmmww)
```

---

## 6. Fractional Ridge Regression (`fracregridge`)

Standard maximum likelihood approaches can suffer from multicollinearity or overfitting when the number of covariates is large relative to the sample size. `fracregridge` implements Fractional Ridge Regression, which applies L2 regularization to shrink coefficients. 

Unlike standard ridge models that require you to specify the penalty term $\alpha$ directly, `fracregridge` allows you to specify a sequence of desired *fractions* representing the target length of the regularised coefficient vector relative to the unregularised OLS vector. The algorithm dynamically calculates the optimal penalty for each fraction.

### 6.1 Shrinking Empirical 401(k) Estimates

Here, we apply fractional ridge regression to the 401(k) data, shrinking the coefficient vector to 20%, 40%, 60%, 80%, and 100% (unregularised) of its original length.

```{r fracregridge}
data("fracreg_k401k")
y_401k <- fracreg_k401k$prate
X_401k <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,
                totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole)

# Fit fractional ridge regression across multiple target fractions
mod_ridge <- fracregridge(y = y_401k, x = X_401k, fracs = seq(0.2, 1.0, by = 0.2))

# View detailed summary showing the chosen alphas and corresponding coefficients
summary(mod_ridge)

# Compute Average Partial Effects for the ridge regularised coefficients
pe_ridge <- fracregridge.pe(mod_ridge)
summary(pe_ridge)
```

---

## 7. Fractional Multinomial Logit (`fracregmlogit`)

When your response variable consists of multiple fractions that always sum to 1 across different categories (e.g., budget shares, time allocation, or market shares), a univariate model is insufficient. `fracregmlogit` estimates these multivariate bounded outcomes simultaneously.

We use the built-in `fracreg_spending` dataset, which records the budget share allocations of local governments across 6 distinct sectors. The sum of these 6 sectoral fractions for any given government is exactly 1.

```{r fracregmlogit}
# Load the empirical spending data
data("fracreg_spending")

# Define covariates and fractional responses (6 categories)
X_mn <- fracreg_spending[, c("houseval", "popdens", "noleft", "minorityleft", "tot")]
y_mn <- fracreg_spending[, c("governing", "safety", "education", "recreation", "social", "urbanplanning")]

# Fit the Fractional Multinomial Logit model
mn_fit <- fracregmlogit(y_mn, X_mn)
summary(mn_fit)

# Compute discrete Partial Effects for specific covariates
mn_pe <- fracregmlogit.pe(mn_fit, effect = "discrete", varlist = c("noleft", "minorityleft"))
summary(mn_pe)
```

### 7.1 Willingness to Pay (WTP)

In policy or consumer choice applications, you might want to calculate the aggregate effect of a variable scaled by arbitrary outcome weights (such as cost, budget, or willingness to pay associated with each choice). The `wtp()` function multiplies the average partial effects by these arbitrary values and tests if the aggregate effect differs from zero.

```{r fracregmlogit_wtp}
# Calculate Willingness to Pay for the 'noleft' variable using a hypothetical WTP vector.
# Because there are 6 choices, we provide a WTP vector of length 6.
wtp_est <- wtp(mn_pe, wtp.vec = 1:6, varlist = "noleft")
summary(wtp_est)

# Plot the Willingness to Pay effect across observations to visualize heterogeneity
plot(mn_fit, wtp.vec = 1:6, varlist = "noleft")
```

## Conclusion

With `fracreg`, `fracreghet`, `fracregpd`, `fracregmlogit`, and `fracregridge`, you have a complete, mathematically robust toolkit for modelling fractional responses. Whether you are dealing with boundary inflation, panel data heterogeneity, endogenous covariates, multinomial shares, or multicollinearity, this package provides consistent, reliable, and easily interpretable econometric solutions.

## Acknowledgements

This package builds upon, consolidates, and modernises the fractional regression frameworks originally implemented in the `frm`, `frmhet`, and `frmpd` R packages developed by Joaquim J.S. Ramalho. As those original packages have been deprecated and removed from the active CRAN repository, `fracreg` serves as an actively maintained successor, ensuring these econometric tools remain available to the R community.

Furthermore, we acknowledge James Ji and A. John Woodill, the authors of the `fmlogit` R package on GitHub, whose foundational work on fractional multinomial logit models inspired the implementation of `fracregmlogit`. We also extend our gratitude to Ariel Rokem and Kendrick Kay, the authors of the `fracridge` package, whose methodological contributions to fractional ridge regression are incorporated into the `fracregridge` functionalities of this package.

## References

- **Ji, J., and Woodill, A. J.** *fmlogit: Fractional Multinomial Logit*. R package repository. <https://github.com/f1kidd/fmlogit>
- **Rokem, A., and Kay, K.** *fracridge: Fractional Ridge Regression*. Package repository. <https://github.com/nrdg/fracridge>

- **Ramalho, J. J. S. (2022).** *frm: Fractional Regression Models*. R package. Formerly available on CRAN, currently archived.
- **Ramalho, J. J. S. (2023).** *frmhet: Fractional Regression Models under Heterogeneity*. R package. Formerly available on CRAN, currently archived.
- **Ramalho, J. J. S. (2023).** *frmpd: Fractional Regression Models for Panel Data*. R package. Formerly available on CRAN, currently archived.
- **Buis, M. L. (2008).** "fmlogit: Stata module fitting a fractional multinomial logit model by quasi maximum likelihood", *Statistical Software Components*, Boston College Department of Economics.
- **Davidson, R. and MacKinnon, J.G. (1981).** "Several tests for model specification in the presence of alternative hypotheses", *Econometrica*, 49(3), 781-793.
- **Fang, K., & Ma, S. (2013).** "Three-part model for fractional response variables with application to Chinese household health insurance coverage", *Journal of Applied Statistics*, 40(5), 925-940.
- **Mullahy, J. (2015).** "Multivariate fractional regression estimation of econometric share models", *Journal of Econometric Methods*, 4(1), 71-100.
- **Murteira, J. M. R., and Ramalho, J. J. S. (2016).** "Regression analysis of multivariate fractional data", *Econometric Reviews*, 35(4), 515-552.
- **Papke, L. E. and Wooldridge, J. M. (1996).** "Econometric methods for fractional response variables with an application to 401(k) plan participation rates", *Journal of Applied Econometrics*, 11(6), 619-632.
- **Papke, L. and Wooldridge, J.M. (2008).** "Panel data methods for fractional response variables with an application to test pass rates", *Journal of Econometrics*, 145(1-2), 121-233.
- **Pregibon, D. (1980).** "Goodness of Link Tests for Generalized Linear Models", *Journal of the Royal Statistical Society: Series C (Applied Statistics)*, 29(1), 15-24.
- **Ramalho, E. A., & Ramalho, J. J. S. (2017).** "Moment-based estimation of nonlinear regression models with boundary outcomes and endogeneity, with applications to nonnegative and fractional responses", *Econometric Reviews*, 36(4), 397-420.
- **Ramalho, E.A., Ramalho, J.J.S. and Murteira, J.M.R. (2011).** "Alternative estimating and testing empirical strategies for fractional response models", *Journal of Economic Surveys*, 25(1), 19-68.
- **Ramalho, E.A., Ramalho, J.J.S. and Murteira, J.M.R. (2014).** "A generalized goodness-of-functional form test for binary and fractional response models", *Manchester School*, 82(4), 488-507.
- **Ramsey, J.B. (1969).** "Tests for Specification Errors in Classical Linear Least-Squares Regression Analysis", *Journal of the Royal Statistical Society: Series B (Methodological)*, 31(2), 350-371.
- **Rokem, A., and Kay, K. (2020).** "Fractional ridge regression: a fast, interpretable reparameterization of ridge regression", *GigaScience*, 9(12), giaa133.

For more information, please visit the [package website](https://sulmanolieko.github.io/fracreg/) or file an issue on [GitHub](https://github.com/SulmanOlieko/fracreg/issues).

To cite this package in your research:

```r
citation("fracreg")
```
