---
title: "A Comprehensive Guide to Fractional Response Regressions with fracreg"
author: "Sulman Olieko Owili"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Comprehensive Guide to Fractional Response Regressions with fracreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = TRUE
)

local({
  hook_output <- knitr::knit_hooks$get("output")
  knitr::knit_hooks$set(output = function(x, options) {
    paste0("\n<details><summary>Toggle to see the output</summary>\n\n",
           hook_output(x, options),
           "\n</details>\n")
  })
})

library(fracreg)
```

## Introduction

The **`fracreg`** package is the most comprehensive tool available in R for estimating, diagnosing, and interpreting fractional response models. It supports:

- **One-part models (1P)**: A single process governs the entire distribution of $y \in [0, 1]$.
- **Two-part hurdle models (2P)**: Separately models the binary hurdle ($y > 0$) and the continuous fractional part ($0 < y \le 1$).
- **Three-part double-inflated models (3P)**: Handles inflation at **both** 0 and 1.
- **Panel data models** (`fracregpd`): CRE, QML, and GMM approaches for longitudinal data.
- **Endogeneity correction** (`fracreghet`): Control function and IV-GMM approaches.

This vignette walks you through comprehensive empirical examples (using the 401(k) dataset) and simulated examples for each estimator.

```{r load-package}
library(fracreg)
```

## 1. Data Description and Preparation

We use the built-in `fracreg_k401k` dataset, which is the canonical firm-level 401(k) plan participation data used in **Papke and Wooldridge (1996)** (*"Econometric methods for fractional response variables with an application to 401(k) plan participation rates"*, Journal of Applied Econometrics). The dataset contains 1,534 observations of 401(k) plans.

The primary variables we use are:
- `prate`: The plan participation rate (the fraction of eligible employees who are active participants). It strictly falls in the $[0, 1]$ interval and is our dependent variable ($y$). 
- `mrate`: The firm's matching rate (the firm's contribution per $1 of employee contribution). 
- `age`: The age of the 401(k) plan. 
- `totemp`: Total number of employees at the firm. 
- `sole`: A binary indicator equal to 1 if the 401(k) plan is the sole retirement plan offered by the firm.

---

## 2. Cross-Sectional Fractional Models (`fracreg`)

The core `fracreg()` function is designed for univariate models where the dependent variable is bounded between 0 and 1 inclusive ($0 \le y \le 1$). The examples below demonstrate both empirical 401(k) plan participation data and general simulated boundaries.

```{r fracreg}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
# 1P Model 
fit <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(fit)

# 1P Model reporting odds ratios and 99% confidence intervals
fit <- fracreg(y, X, type="1P", linkfrac="logit", or=TRUE, level=0.99)
summary(fit)
 
# 2P Model (modelling mass at 1) 
fit <- fracreg(y, X, type="2P", inflation=1, linkbin="logit", linkfrac="logit") 
summary(fit)
 
# 3P Model (inject artificial 0s for demonstration) 
y_3p <- y; y_3p[1:50] <- 0 
fit <- fracreg(y_3p, X, type="3P", linkbin=c("logit","logit"), linkfrac="logit") 
summary(fit)
 
### Simulated Examples 
 
set.seed(123) 
N <- 1000 
x1 <- rnorm(N) 
x2 <- runif(N) 
 
# Generating a fractional dependent variable with inflation at 0 and 1 
XB <- -0.5 + 0.8 * x1 + 1.2 * x2 + rnorm(N) 
y_latent <- exp(XB) / (1 + exp(XB)) 
 
y <- y_latent 
# Inflate at boundaries 
y[y_latent < 0.2] <- 0 
y[y_latent > 0.8] <- 1 
 
X <- cbind(x1 = x1, x2 = x2) 
 
# fracreg estimation of a logit fractional response model 
fit <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(fit)
 
# fracreg estimation of the binary logit component of the two-part fractional 
# regression model with y=0 as the relevant boundary value 
fit <- fracreg(y, X, type="2Pbin", inflation=0, linkbin="logit") 
summary(fit)
 
# fracreg estimation of the fractional component of the two-part fractional 
# regression model with y=0 as the relevant boundary value and using a 
# probit link function 
fit <- fracreg(y, X, type="2Pfrac", inflation=0, linkfrac="probit") 
summary(fit)
 
# fracreg estimation of both components of a two-part fractional response model 
# with y=0 as the relevant boundary value and using a cloglog binary link 
# function and a logit fractional link function 
fit <- fracreg(y, X, type="2P", inflation=0, linkbin="cloglog", linkfrac="logit") 
summary(fit)
 
# Three-part double-inflated model (y has both 0s and 1s) 
fit <- fracreg(y, X, type="3P", linkbin=c("logit","probit"), linkfrac="logit") 
summary(fit)
```

### 1.1 Partial Effects (`fracreg.pe`)

Raw coefficients from fractional models are not directly interpretable as marginal effects. Use `fracreg.pe()` to compute the Average Partial Effects (APE) and Conditional Partial Effects (CPE) using the analytical delta method.

```{r fracreg-pe}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
m <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(m)
fracreg.pe(m) 
 
### Simulated Examples 
 
N <- 250 
u <- rnorm(N) 
 
X <- cbind(rnorm(N),rnorm(N)) 
dimnames(X)[[2]] <- c("X1","X2") 
 
ym <- exp(X[,1]+X[,2]+u)/(1+exp(X[,1]+X[,2]+u)) 
y <- rbeta(N,ym*20,20*(1-ym)) 
y[y > 0.9] <- 1 
 
#Computing average partial effects for a logit fractional response model 
res <- fracreg(y,X,linkfrac="logit",table=FALSE) 
summary(res)
fracreg.pe(res) 
 
#Computing average partial effects for a binary logit + fractional probit 
#two-part model 
res <- fracreg(y,X,linkbin="logit",linkfrac="probit",type="2P",inf=1,table=FALSE) 
summary(res)
fracreg.pe(res) 
 
#Computing conditional partial effects for X2 in the logit component 
#of a two-part fractional response model, with the covariates evaluated 
#at their median values 
res <- fracreg(y,X,linkfrac="logit",type="2Pfrac",inf=1,table=FALSE) 
summary(res)
fracreg.pe(res,APE=FALSE,CPE=TRUE,at="median",which.x="X2") 
 
#Computing average partial effects for a three-part double-inflated model 
y3p <- y 
y3p[1:20] <- 0 
y3p[21:40] <- 1 
res3p <- fracreg(y3p,X,linkbin=c("logit","probit"),linkfrac="logit",type="3P",table=FALSE) 
summary(res3p)
fracreg.pe(res3p) 
```

---

## 3. Hypothesis Testing and Specification Diagnostics

`fracreg` includes state-of-the-art specification tests to validate your model's functional form and link function assumptions.

### 3.1 Generalised Goodness-Of-Functional-Form (`fracreg.ggoff`)

The GGOFF test tests whether the chosen link function is adequate for the data. A significant result suggests the link may be misspecified.

```{r fracreg-ggoff}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
m <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(m)
fracreg.ggoff(m) 
 
### Simulated Examples 
 
N <- 250 
u <- rnorm(N) 
 
X <- cbind(rnorm(N),rnorm(N)) 
dimnames(X)[[2]] <- c("X1","X2") 
 
ym <- exp(X[,1]+X[,2]+u)/(1+exp(X[,1]+X[,2]+u)) 
y <- rbeta(N,ym*20,20*(1-ym)) 
y[y > 0.9] <- 1 
 
#Testing the logit specification of a standard fractional response model 
#using LM and Wald versions of the GGOFF test, based on 1 or 2 fitted powers of 
#the linear predictor 
res <- fracreg(y,X,linkfrac="logit",table=FALSE) 
summary(res)
fracreg.ggoff(res,c("Wald","LM")) 
 
#Testing the probit specification of the binary component of a two-part fractional 
#regression model using a LR-based GGOFF test 
res <- fracreg(y,X,linkbin="probit",type="2Pbin",inf=1,table=FALSE) 
summary(res)
fracreg.ggoff(res,"LR") 
```

### 3.2 RESET Test (`fracreg.reset`)

The RESET test detects general functional form misspecification by testing whether powers of the fitted values have explanatory power. Testing $H_0: \gamma = 0$ provides a robust diagnostic check.

```{r fracreg-reset}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
m <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(m)
fracreg.reset(m) 
 
### Simulated Examples 
 
N <- 250 
u <- rnorm(N) 
 
X <- cbind(rnorm(N),rnorm(N)) 
dimnames(X)[[2]] <- c("X1","X2") 
 
ym <- exp(X[,1]+X[,2]+u)/(1+exp(X[,1]+X[,2]+u)) 
y <- rbeta(N,ym*20,20*(1-ym)) 
y[y > 0.9] <- 1 
 
#Testing the logit specification of a standard fractional response model 
#using LM and Wald versions of the RESET test, based on 1 or 2 fitted powers of 
#the linear predictor 
res <- fracreg(y,X,linkfrac="logit",table=FALSE) 
summary(res)
fracreg.reset(res,2:3,c("Wald","LM")) 
 
#Testing the probit specification of the binary component of a two-part fractional 
#regression model using LR-based RESET tests with quadratic and cubic fitted  
#powers of the linear predictor 
res <- fracreg(y,X,linkbin="probit",type="2Pbin",inf=1,table=FALSE) 
summary(res)
fracreg.reset(res,3,"LR") 
```

### 3.3 P-Test for Non-Nested Models (`fracreg.ptest`)

You can compare non-nested models (e.g., `logit` vs. `cloglog` link) using the Davidson-MacKinnon (1981) P-test.

```{r fracreg-ptest}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
m1 <- fracreg(y, X, type="1P", linkfrac="logit") 
summary(m1)
m2 <- fracreg(y, X, type="1P", linkfrac="probit") 
summary(m2)
fracreg.ptest(m1, m2) 
 
### Simulated Examples 
 
N <- 250 
u <- rnorm(N) 
 
X <- cbind(rnorm(N),rnorm(N)) 
dimnames(X)[[2]] <- c("X1","X2") 
 
ym <- exp(X[,1]+X[,2]+u)/(1+exp(X[,1]+X[,2]+u)) 
y <- rbeta(N,ym*20,20*(1-ym)) 
y[y > 0.9] <- 1 
 
#Testing logit versus loglog specifications for standard fractional 
#regression models using a LM version of the P test 
res1 <- fracreg(y,X,linkfrac="logit",table=FALSE) 
summary(res1)
res2 <- fracreg(y,X,linkfrac="loglog",table=FALSE) 
summary(res2)
fracreg.ptest(res1,res2,"LM") 
 
#Testing a logit one-part fractional response model versus a binary logit + 
#fractional probit two-part model using a Wald version of the P test 
res1 <- fracreg(y,X,linkfrac="logit",table=FALSE) 
summary(res1)
res2 <- fracreg(y,X,linkbin="logit",linkfrac="probit",type="2P",inf=1,table=FALSE) 
summary(res2)
fracreg.ptest(res1,res2,"Wald") 
```

---

## 4. Endogeneity & Heteroscedasticity (`fracreghet`)

When you suspect that one of your covariates is endogenous, or that the variance is heteroscedastic, `fracreghet()` provides instrumental variable correction. It natively supports IV through a **Control Function (CF)** approach or **GMM** estimation.

```{r fracreghet}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X_het <- cbind(mrate = fracreg_k401k$mrate, ltotemp = fracreg_k401k$ltotemp)
 
# fracreghet estimators do not allow exact 1s or 0s
y_adj <- y
y_adj[y_adj == 1] <- 0.999

# Instrument mrate using age

Z_emp <- cbind(age = fracreg_k401k$age, ltotemp = fracreg_k401k$ltotemp) 
fit <- fracreghet(y_adj, X_het, Z_emp, var.endog = X_het[, "mrate"], type="QMLxv", link="logit") 
summary(fit)

# Compute the same QMLxv estimator reporting Odds Ratios with 90% confidence intervals
fit <- fracreghet(y_adj, X_het, Z_emp, var.endog = X_het[, "mrate"], type="QMLxv", link="logit", or=TRUE, level=0.90)
summary(fit)
 
### Simulated Examples 
 
set.seed(123) 
N <- 1000 
x1 <- rnorm(N) 
 
# Simulating an endogenous variable (var.endog) and an instrument (z1) 
z1 <- rnorm(N) 
u <- 0.5 * z1 + rnorm(N) 
var.endog <- 0.8 * z1 + u 
y_endog <- exp(0.5 * x1 + 1.2 * var.endog + u) / (1 + exp(0.5 * x1 + 1.2 * var.endog + u)) 
 
# Avoid exact 0 or 1 boundaries for some estimators 
y_endog[y_endog <= 0] <- 0.01 
y_endog[y_endog >= 1] <- 0.99 
 
X <- cbind(x1 = x1, var.endog = var.endog) 
Z <- cbind(x1 = x1, z1 = z1) 
 
# Exogeneity (assuming var.endog is exogenous for comparison), GMMx estimator 
fit <- fracreghet(y = y_endog, x = X, type = "GMMx", link = "logit") 
summary(fit)
 
# Endogeneity, GMMz estimator (does not require reduced form for endog) 
fit <- fracreghet(y = y_endog, x = X, z = Z, type = "GMMz", link = "logit") 
summary(fit)
 
# Endogeneity, GMMxv estimator (assumes linear reduced form for var.endog) 
fit <- fracreghet(y = y_endog, x = X, z = Z, var.endog = var.endog, type = "GMMxv", link = "logit") 
summary(fit)
 
# Endogeneity, QMLxv control function approach 
fit <- fracreghet(y = y_endog, x = X, z = Z, var.endog = var.endog, type = "QMLxv", link = "logit") 
summary(fit)
```

### 4.1 Partial Effects for Endogenous Models (`fracreghet.pe`)

```{r fracreghet-pe}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X_het <- cbind(mrate = fracreg_k401k$mrate, ltotemp = fracreg_k401k$ltotemp)
 
# fracreghet estimators do not allow exact 1s or 0s
y_adj <- y
y_adj[y_adj == 1] <- 0.999

# Instrument mrate using age

Z_emp <- cbind(age = fracreg_k401k$age, ltotemp = fracreg_k401k$ltotemp) 
res_emp <- fracreghet(y_adj, X_het, Z_emp, var.endog = X_het[, "mrate"], type="QMLxv", link="logit", table=FALSE) 
summary(res_emp)
fracreghet.pe(res_emp, which.x="mrate")

### Simulated Examples 
 
N <- 250 
u <- rnorm(N) 
 
X <- cbind(rnorm(N),rnorm(N)) 
dimnames(X)[[2]] <- c("X1","X2") 
 
Z <- cbind(rnorm(N),rnorm(N),rnorm(N)) 
dimnames(Z)[[2]] <- c("Z1","Z2","Z3") 
 
y <- exp(X[,1]+X[,2]+u)/(1+exp(X[,1]+X[,2]+u)) 
 
res <- fracreghet(y,X,type="GMMx",table=FALSE) 
summary(res)
 
#Smearing estimator of average partial effects for variable X1 
fracreghet.pe(res,which.x="X1") 
 
#Naive estimator of conditional partial effects for all covariates, 
#which are evaluated at X1=1 and X2=-1 
fracreghet.pe(res,smearing=FALSE,APE=FALSE,CPE=TRUE,at=c(1,-1)) 
```

### 4.2 RESET Test for Endogenous Models (`fracreghet.reset`)

```{r fracreghet-reset}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X_het <- cbind(mrate = fracreg_k401k$mrate, ltotemp = fracreg_k401k$ltotemp)
 
# fracreghet estimators do not allow exact 1s or 0s
y_adj <- y
y_adj[y_adj == 1] <- 0.999

# Instrument mrate using age

Z_emp <- cbind(age = fracreg_k401k$age, ltotemp = fracreg_k401k$ltotemp) 
res_emp <- fracreghet(y_adj, X_het, type="GMMx", link="logit", table=FALSE) 
summary(res_emp)
fracreghet.reset(res_emp)

### Simulated Examples 
 
N <- 250 
u <- rnorm(N) 
 
X <- cbind(rnorm(N),rnorm(N)) 
dimnames(X)[[2]] <- c("X1","X2") 
 
Z <- cbind(rnorm(N),rnorm(N),rnorm(N)) 
dimnames(Z)[[2]] <- c("Z1","Z2","Z3") 
 
y <- exp(X[,1]+X[,2]+u)/(1+exp(X[,1]+X[,2]+u)) 
 
res <- fracreghet(y,X,type="GMMx",table=FALSE) 
summary(res)
 
#LM and Wald versions of the RESET test, based on 1 or 2 fitted powers of xb 
fracreghet.reset(res,2:3,c("Wald","LM")) 
```

---

## 5. Panel Data Fractional Models (`fracregpd`)

For longitudinal or panel data where unobserved heterogeneity is a concern, `fracregpd()` provides fixed-T panel estimators. 

```{r fracregpd}
### Empirical 401(k) Examples 
data("fracreg_k401k") 
y <- fracreg_k401k$prate 
X <- cbind(mrate = fracreg_k401k$mrate, age = fracreg_k401k$age,  
           totemp = fracreg_k401k$totemp, sole = fracreg_k401k$sole) 
 
# Artificial panel data structure for demonstration 
N_emp <- nrow(X) 
id_emp <- rep(1:(N_emp/2), each=2) 
time_emp <- rep(1:2, times=N_emp/2) 
fit <- fracregpd(id_emp, time_emp, y, X, type="QMLcre", link="probit") 
summary(fit)
 
### Simulated Examples 
 
set.seed(123) 
# Simulating Panel Data 
N <- 100 
T_periods <- 5 
id <- rep(1:N, each = T_periods) 
time <- rep(1:T_periods, times = N) 
x_panel <- rnorm(N * T_periods) 
 
# Unobserved individual effect (CRE) 
c_i <- rep(rnorm(N), each = T_periods)  
y_panel <- exp(x_panel + c_i) / (1 + exp(x_panel + c_i)) 
 
X <- cbind(x_panel = x_panel) 
 
# Endogenous variable and instrument simulation 
z_panel <- rnorm(N * T_periods) 
u_panel <- 0.5 * z_panel + rnorm(N * T_periods) 
var_endog <- 0.8 * z_panel + u_panel 
y_endog <- exp(x_panel + 1.2 * var_endog + c_i + u_panel) /  
             (1 + exp(x_panel + 1.2 * var_endog + c_i + u_panel)) 
 
X_endog <- cbind(x_panel = x_panel, var_endog = var_endog) 
Z_inst <- cbind(x_panel = x_panel, z_panel = z_panel) 
 
# Estimate a Correlated Random Effects (CRE) Model 
fit <- fracregpd(id=id, time=time, y=y_panel, x=X, type="QMLcre", link="probit") 
summary(fit)
 
# Exogeneity, no lags, no time dummies, clustered standard errors, GMMbgw estimator 
fit <- fracregpd(id=id, time=time, y=y_panel, x=X, type="GMMbgw") 
summary(fit)

# Estimate the GMMww estimator with odds ratios and 99% confidence intervals
fit <- fracregpd(id=id, time=time, y=y_panel, x=X, type="GMMww", or=TRUE, level=0.99)
summary(fit)
 
# Lagged covariates and instruments, robust standard errors, GMMww estimator 
fit <- fracregpd(id=id, time=time, y=y_panel, x=X, lags=TRUE, type="GMMww", var.type="robust") 
summary(fit)
 
# Endogeneity, time dummies, GMMpfe estimator 
fit <- fracregpd(id=id, time=time, y=y_endog, x=X_endog, z=Z_inst,  
          x.exogenous=FALSE, type="GMMpfe", tdummies=TRUE) 
summary(fit)
```

---

## Conclusion

With `fracreg`, `fracreghet`, and `fracregpd`, you have a complete, mathematically robust toolkit for modelling fractional responses bounded between $[0,1]$, regardless of inflation, endogeneity, or unobserved panel effects.

## Acknowledgements

This package builds upon, consolidates, and modernises the fractional
regression frameworks originally implemented in the `frm`, `frmhet`, and
`frmpd` R packages developed by Joaquim J.S. Ramalho. As those original
packages have been deprecated and removed from the active CRAN
repository, `fracreg` serves as an actively maintained successor,
ensuring these econometric tools remain available to the R community.

## References

- **Ramalho, J. J. S. (2022).** *frm: Fractional Regression Models*. R
  package. Formerly available on CRAN, currently archived.
- **Ramalho, J. J. S. (2023).** *frmhet: Fractional Regression Models
  under Heterogeneity*. R package. Formerly available on CRAN, currently
  archived.
- **Ramalho, J. J. S. (2023).** *frmpd: Fractional Regression Models for
  Panel Data*. R package. Formerly available on CRAN, currently
  archived.
- **Davidson, R. and MacKinnon, J.G. (1981).** "Several tests for model specification in the presence of alternative hypotheses", *Econometrica*, 49(3), 781-793.
- **Fang, K., & Ma, S. (2013).** "Three-part model for fractional response variables with application to Chinese household health insurance coverage", *Journal of Applied Statistics*, 40(5), 925-940.
- **Papke, L. E. and Wooldridge, J. M. (1996).** "Econometric methods for fractional response variables with an application to 401(k) plan participation rates", *Journal of Applied Econometrics*, 11(6), 619-632.
- **Papke, L. and Wooldridge, J.M. (2008).** "Panel data methods for fractional response variables with an application to test pass rates", *Journal of Econometrics*, 145(1-2), 121-233.
- **Pregibon, D. (1980).** "Goodness of Link Tests for Generalized Linear Models", *Journal of the Royal Statistical Society: Series C (Applied Statistics)*, 29(1), 15-24.
- **Ramalho, E. A., & Ramalho, J. J. S. (2017).** "Moment-based estimation of nonlinear regression models with boundary outcomes and endogeneity, with applications to nonnegative and fractional responses", *Econometric Reviews*, 36(4), 397-420.
- **Ramalho, E.A., Ramalho, J.J.S. and Murteira, J.M.R. (2011).** "Alternative estimating and testing empirical strategies for fractional response models", *Journal of Economic Surveys*, 25(1), 19-68.
- **Ramalho, E.A., Ramalho, J.J.S. and Murteira, J.M.R. (2014).** "A generalized goodness-of-functional form test for binary and fractional response models", *Manchester School*, 82(4), 488-507.
- **Ramsey, J.B. (1969).** "Tests for Specification Errors in Classical Linear Least-Squares Regression Analysis", *Journal of the Royal Statistical Society: Series B (Methodological)*, 31(2), 350-371.

For more information, please visit the [package website](https://sulmanolieko.github.io/fracreg/) or file an issue on [GitHub](https://github.com/SulmanOlieko/fracreg/issues).

To cite this package in your research:

```r
citation("fracreg")
```
